\chapter*{List of publications}
\addcontentsline{toc}{part}{List of publications}

\section*{Published}
\begin{itemize}
\item Doury M, Dizeux A, De Cesare A, Lucidarme O, Bridal LS, Frouin F, Regularized linear resolution of a one-compartment model to improve the reproducibility of perfusion parameters in CEUS, IEEE International Ultrasonics Symposium (IUS), Tours, France, 2016.
\item Doury M, Dizeux A, De Cesare A, Lucidarme O, Pellot-Barakat C, Bridal LS, Frouin F, Quantification of tumor perfusion using dynamic contrast-enhanced ultrasound: Impact of Mathematical Modeling, Physics in Medicine and Biology, 2017.
\end{itemize}

\section*{In press}
\begin{itemize}
\item Doury M, De Cesare A, Bridal LS, Pellot-Barakat C, Frouin F, Impact of Recirculation in Dynamic Contrast-Enhanced Ultrasound: a Simulation Study, IRBM, 2017.
\end{itemize}

\section*{In writing}
\begin{itemize}
\item Doury M, De Cesare A, Pellot-Barakat C, Bridal LS, Frouin F, Error Sources Affecting Relative Quantification of Dynamic Contrast-Enhanced Ultrasound, Medical Image Analysis
\end{itemize}

\section*{Oral communications}
\begin{itemize}
\item Doury M, Comparison of three modelling approaches in CEUS studies, Dynamics and Control of Tumor Growth Workshop (DYCOTUG), Rouen, France, 2015.
\item Doury M, Dizeux A, De Cesare A, Lucidarme O, Bridal SL, Frouin F, Regularized linear resolution of a one-compartment model to improve the reproducibility of perfusion parameters in CEUS, IEEE International Ultrasonics Symposium (IUS), Tours, France, 2016.
\item Doury M, Bridal LS, Frouin F, How to measure tumor perfusion for exam comparison?, RITS 2017, Lyon, France, 2017.
\end{itemize}

\section*{Posters}
\begin{itemize}
\item Doury M, Dizeux A, Barrois G, Lucidarme O, Bridal L, De Cesare A, Frouin F, Suivi longitudinal en imagerie ultrasonore multimodale pour la caract\'erisation d’un mod\`ele tumoral, Journ\'ees de l’\'Ecole Doctorale Pierre Louis de Sant\'e Publique (ED393), Saint-Malo, France, 2014.
\item Doury M, Dizeux A, Barrois G, Le Guillou-Buffelo D, Coron A, Bridal L, De Cesare A, Frouin F, Local classification of microvascular function based on contrast-enhanced ultrasound data: a feasibility study, Journ\'ees de Recherche en Imagerie et Technologies pour la Sant\'e (RITS), Dourdan, France, 2015.
\item Doury M, Dizeux A, Bridal L, De Cesare A, Frouin F, Impact of three modeling approaches on the reproducibility of perfusion parameters in CEUS studies, Journ\'ees de l’\'Ecole Doctorale Pierre Louis de Sant\'e Publique (ED393), Saint-Malo, France, 2015.
\item Doury M, Dizeux A, De Cesare A, Lucidarme O, Bridal SL, Frouin F, Comparison of three modelling approaches on the reproducibility of perfusion parameters in CEUS studies, International Symposium on Biomedical Imaging: From Nano to Macro (ISBI), Prague, Czech Republic, 2016.
\item Doury M, Dizeux A, De Cesare A, Lucidarme O, Bridal SL, Frouin F, Impact of the modelling approach on the reproducibility of perfusion parameters in CEUS studies, Journ\'ees de l’\'Ecole Doctorale Pierre Louis de Sant\'e Publique (ED393), Saint-Malo, France, 2016.
\item Doury M, De Cesare A, Bridal LS, Frouin F, Impact of Recirculation in Dynamic Contrast Enhanced Ultrasound: A Simulation Study, RITS, Lyon, France, 2017. Prix jeune chercheur SFGBM.
\end{itemize}